---
title: "Immuno Metabolome Atlas"
output:
  github_document:
    df_print: kable
---

```{r, echo=F, warning=FALSE, message = FALSE}
options(warn=-1)
library(ImmunoMet, warn.conflicts=FALSE, quietly = TRUE)
load_data("C:/Users/Pascal/Documents/ImmuneMetAtlas/App/config.yaml", neighbours = 0)
```


# Introduction
This package is part of our publication called "Immuno metabolome atlas". 

# Installation

##  Python

Dependencies needed:

- Pandas
- Numpy

The script will attempt to install these if they cannot be imported. 


## R
Use the `devtools` to install our package and dependencies.

``` {r eval=FALSE}
install.packages("devtools")
devtools::install_github("vanhasseltlab/ImmuneMetAtlas")
```

# Getting Started

## Setting up the configuration file
To link the Python and R scripts, a YAML (Yet Another Markdown Language) configuration file is used, called `config.yaml`. This file needs to contain at least the follwing fields:

- folder: Path where data files are going to be stored
- GO_ID: Gene Ontology identifier of which all offspring terms are used as a filter


## Preprocessing data
Before using this package, you will need to obtain data using our Python scripts. These are located in the Python folder and only has be executed once. Preprocessing is done by running `Preprocessing.py`, which can be done by either of the following options:

- Retain the files in the current folder structure and run the script (default)
- Change the variable `config_path` in `Preprocessing.py` to its actual location
- Run the script in a terminal with the first argument being the path to `config.yaml` as follows:

```{bash eval=FALSE}
python3 Preprocessing.py path_to_config.yaml
```

The scripts will download and extract all data to the given folder in the configuration file. 

## Start the RShiny app

``` {r eval=FALSE}
library("ImmunoMet")
load_data(config = "config.yaml")
run_shiny()
```

## Non-Shiny usages

Example, static graph
``` {r eval=FALSE}
library("ImmunoMet")
load_data(config = "config.yaml")
graph <- example_graph()
plot(graph)
```

```{r results='hide', fig.height=6, echo=F, fig.align='center'}
graph <- example_graph()
plot(graph)
```

Interactive plotly graph in browser

``` {r eval=FALSE}
to_plotly(graph)
```



Basically `get_graph()` is the main function for obtaining a graph-structure representing your search. It takes the parameters `filter`, `type`, `search_mode`, `neighbours`, `max_neighbours` and `simple`. See the documentation within the R package for full explanation and examples of these parameters.

```{r eval=FALSE}
graph <- get_graph(filter = "microglial cell activation", type = "Gene Ontology")
plot(graph)
```




Obtain p-values of metabolite - GO associations in a long dataframe format:
``` {r eval=FALSE}
head(get_metabolite_go_pvalues(graph))
```

When you only need the graph structure, use `simple = TRUE`. This is especially handy when constructing a lot of graphs. Note that only the bare minimum data is stored. 

``` {r}
get_graph(filter = "microglial cell activation", type = "Gene Ontology", simple = TRUE)
```